---
title: "USGS Flow (Q) Gages in CA by County"
format:
  #pdf: default
  html:
    theme: materia # sandstone, materia
    toc: true
    toc-depth: 2
    self-contained: true
    code-fold: true
    code-overflow: wrap
format-links: false
execute:
  echo: false
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: load
#| message: false
targets::tar_load(ca_Q_sites_cnty)
targets::tar_load(ca_counties)
library(tidyverse)
library(sf)

```

A quick synopsis of USGS flow (discharge "**Q**") stations in CA by county. 

## Methods

A current list of USGS flow gages in California was downloaded via the {dataRetrieval} package in R on `r Sys.Date()`. These data were then filtered to sites that contain valid latitude and longitude data, and are spatially mapped and summarized below.

## Summary across CA

```{r}
#| label: fig-ca-map
#| fig-cap: "Map of USGS gages collecting discharge (Q) in California."

ggplot() + 
  geom_sf(data=ca_counties, color="gray20", fill=alpha("gray70",0.2), lwd=0.2)+
  geom_sf(data=ca_Q_sites_cnty, pch=16, color=alpha("steelblue", 0.5)) +
  ggthemes::theme_map()

```

There are a total of `r nrow(ca_Q_sites_cnty)` in California, however these sites are not evenly distributed throughout the state (@fig-ca-map).

## By County

If we summarize the total number of gages in each county, we can see well gaged and poorly gaged counties more clearly (@fig-gage-by-county). These counties are not evenly distributed, but the southern half of CA has a large majority of the USGS flow gages in CA (@fig-county-topbottom).

```{r}
#| label: fig-gage-by-county
#| fig-cap: "Gages by County"

# drop geom and get tally
ca_Q_sites_tally <- ca_Q_sites_cnty |> st_drop_geometry() |> 
  group_by(NAME) |> tally() |> arrange(desc(n))

# get top and bottom 10 counties
top10 <- ca_Q_sites_tally |> slice_max(order_by = n, n=10) |> 
  mutate(group="Top 10")
bot10 <- ca_Q_sites_tally |> slice_min(order_by = n, n=10) |> 
    mutate(group="Bottom 10")

ca_Q_tally_combined <- bind_rows(top10, bot10) 

# barplot
ggplot() + 
  geom_col(data=ca_Q_tally_combined, aes(y=forcats::fct_reorder(NAME, n), x=n)) +
  facet_wrap(~group, drop = TRUE, scales = "free") +
  cowplot::theme_minimal_vgrid() +
  labs(x="Total Q Gages", y="")

```

```{r}
#| label: fig-county-topbottom
#| fig-cap: "Top and bottom 10 counties for number of USGS gages."

ca_Q_sites_cnty |> right_join(ca_Q_tally_combined) |> 
  ggplot() + 
  geom_sf(data=ca_counties, color="gray20", fill=alpha("gray70",0.1), lwd=0.2)+
  geom_sf(data=ca_counties |> filter(NAME %in% ca_Q_tally_combined$NAME), color="gray20", fill=alpha("gray40",0.4), lwd=0.3)+
  geom_sf(aes(fill=group), pch=21, color=alpha("gray70", 0.2)) +
  scale_fill_viridis_d("Top or Bottom", alpha=0.6)+
  cowplot::theme_map()

```

