---
title: "USGS Flow (Q) Gages in CA by County"
format:
  #pdf: default
  html:
    theme: materia # sandstone, materia
    toc: true
    toc-depth: 2
    self-contained: true
    code-fold: true
    code-overflow: wrap
format-links: false
execute:
  echo: false
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: load
#| message: false
targets::tar_load(ca_Q_sites_cnty)
targets::tar_load(ca_counties)
library(tidyverse)
library(sf)
library(glue)
library(ggthemes)

```

A quick synopsis of USGS flow (discharge "**Q**") stations in CA by county. 

## Methods

A current list of USGS flow gages in California was downloaded via the {dataRetrieval} package in R on `r Sys.Date()`. These data were then filtered to sites that contain valid latitude and longitude data, and are spatially mapped and summarized below.

## Summary across CA

```{r}
#| label: fig-ca-map
#| fig-cap: "Map of USGS gages collecting discharge (Q) in California."

ggplot() + 
  geom_sf(data=ca_counties, color="gray20", fill=alpha("gray70",0.2), lwd=0.2)+
  geom_sf(data=ca_Q_sites_cnty, pch=16, color=alpha("steelblue", 0.5)) +
  ggthemes::theme_map()

```

There are a total of `r nrow(ca_Q_sites_cnty)` in California, however these sites are not evenly distributed throughout the state or through the `r nrow(ca_counties)` counties in California (@fig-ca-map).

## By County

If we summarize the total number of gages in each county, we can see well gaged and poorly gaged counties more clearly (@fig-gage-by-county). These well-gaged or not poorly-gaged counties are not evenly distributed. The southern half of CA has a majority of the USGS flow gages in CA (@fig-county-topbottom).

```{r}
#| label: fig-gage-by-county
#| fig-cap: "Gages by County"

# drop geom and get tally
ca_Q_sites_tally <- ca_Q_sites_cnty |> st_drop_geometry() |> 
  group_by(NAME) |> tally() |> arrange(desc(n))

# get top and bottom 10 counties
top10 <- ca_Q_sites_tally |> slice_max(order_by = n, n=10) |> 
  mutate(group="Top 10")
bot10 <- ca_Q_sites_tally |> slice_min(order_by = n, n=10) |> 
    mutate(group="Bottom 10")

ca_Q_tally_combined <- bind_rows(top10, bot10) 

# barplot
ggplot() + 
  geom_col(data=ca_Q_tally_combined, aes(y=forcats::fct_reorder(NAME, n), x=n)) +
  facet_wrap(~group, drop = TRUE, scales = "free") +
  cowplot::theme_minimal_vgrid() +
  labs(x="Total Q Gages", y="")

```

```{r}
#| label: fig-county-topbottom
#| fig-cap: "Top and bottom 10 counties for number of USGS gages."

ca_Q_sites_cnty |> right_join(ca_Q_tally_combined) |> 
  ggplot() + 
  geom_sf(data=ca_counties, color="gray20", fill=alpha("gray70",0.1), lwd=0.2)+
  geom_sf(data=ca_counties |> filter(NAME %in% ca_Q_tally_combined$NAME), color="gray20", fill=alpha("gray40",0.4), lwd=0.3)+
  geom_sf(aes(fill=group), pch=21, color=alpha("gray70", 0.2)) +
  scale_fill_viridis_d("Top or Bottom", alpha=0.6)+
  cowplot::theme_map()

```

```{r}
#| label: example-county
#| echo: false
example_county <- "Mendocino"

```

If we zoom into a single county, it's possible to better assess the distribution of gages within watersheds. Here is an example using `r example_county` County (@fig-county-basemap).

```{r}
#| label: get-basemap
#| echo: false

library(terra)
library(tidyterra)
library(basemaps)

#set_defaults(map_service = "osm", map_type = "topographic")
#set_defaults(map_service = "esri", map_type = "natgeo_world_map")
set_defaults(map_service = "esri", map_type = "world_imagery")

# get basemap and make a spatraster
cnty_rast <- basemap_raster(ca_counties[ca_counties$NAME==example_county,],
                            force = TRUE, verbose = FALSE)
cnty_terra <- terra::rast(cnty_rast) # convert to terra

# mask to county
cnty_terra_mask <- terra::crop(cnty_terra,
                               st_transform(ca_counties[ca_counties$NAME==example_county,], 3857),
                               mask=TRUE)
```

```{r}
#| label: fig-county-basemap
#| fig-cap: "Example county map."

# county map
(gg_cnty_map <-
   ggplot() +
   # cropped only
   geom_spatraster_rgb(data=cnty_terra_mask, alpha=0.8) +
   geom_sf(data=ca_Q_sites_cnty |> filter(NAME==example_county),
           fill="yellow", alpha=0.8, pch=21, 
           col=alpha("gray40", 0.5), size=3) +
   geom_sf(data=ca_counties |> filter(NAME==example_county), 
           fill=alpha("gray70", 0), color="orange", lwd=1) +
   ggspatial::annotation_north_arrow(location="br", 
                                     width = unit(1.2, "cm"), 
                                     pad_y = unit(1.1, "cm")) +
   ggspatial::annotation_scale(unit_category="imperial", location="br") +
   labs(title = glue("USGS gages in {example_county} County"))+
   theme_void()
)
```

